\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-10-07 chapters/bci.tex
funding

\todo{comments liesel paradigms}
\todo{spellcheck}
\todo{are all acronyms defined}
\todo{fix all latex warnings to make sure no refs are missing (or grep on that
specific warning)}
\todo{uniform ref or autoref}
\todo{don't capitalize figure, table, algorithm,...}
\todo{define short captions for tables and figures}
\todo{fix lacheck/chktex notifications}
\todo{use terminology individuals with severe
speech and physical impairments (SSPI), avoid phrasing 'suffering from'}
\todohsm{more subtitles}
\todofc{double check statements about what has been published 'for the sake of clarity, the introduction has been mofed to"}
\todofc{chapters should look more like full papers}
\todo{uniform vsa condition vsa setting}
\todo{acronyms written in full every new chapter and in captions}
\todo{search for ??}
\todo{acro usage in chapters}
\chapter{Brain-computer interfaces}
\label{sec:bci}
People with \ac{sspi} due to
neurodegenenerative disease, stroke, traumatic brain or spine injury
become entirely reliant on caregivers or family members in their day to day
routine.
Wouldn't it be great if we could help them regain some quality of life by
making them more independent?

@@

\subsection{Decoders}
\todo{averaging over trials}
Several techniques are suited to decode~\acp{erp}.
Below are some methods that can achieve state-of-the art performance and are
relevant to this work.

>>>>

\todo{comments liesel paradigms}
\todo{spellcheck}
\todo{are all acronyms defined}
\todo{fix all latex warnings to make sure no refs are missing (or grep on that
specific warning)}
\todo{uniform ref or autoref}
\todo{don't capitalize figure, table, algorithm,...}
\todo{define short captions for tables and figures}
\todo{fix lacheck/chktex notifications}
\todo{use terminology individuals with severe
speech and physical impairments (SSPI), avoid phrasing 'suffering from'}
\todohsm{more subtitles}
\todofc{double check statements about what has been published 'for the sake of clarity, the introduction has been mofed to"}
\todofc{chapters should look more like full papers}
\todo{uniform vsa condition vsa setting}
\todo{acronyms written in full every new chapter and in captions}
\todo{search for ??}
\todo{acro usage in chapters}
\chapter{Brain-computer interfaces}
\todo{are textcites used when necessary?}
\label{sec:bci}
People with \ac{sspi} due to
neurodegenenerative disease, stroke, traumatic brain or spine injury
become entirely reliant on caregivers or family members in their day to day
routine.
Wouldn't it be great if we could help them regain some quality of life by
making them more independent?

@@

\subsection{Decoders}
Due to the unfavorable \ac{snr} of \acp{erp}, it is often not feasible to
accurately decode the encoded information based on just one example.
Therefore, classification algorithm are trained on a set of epochs, but are
often evaluated on an average of multiple epochs.
These averages are constructed over multiple stimulations of the stimulus
representing a given selection, enhancing the \ac{snr}.
The drawback, however, is that performing these multiple stimulations takes
more time.
To optimize the \ac{itr}, the trade-off between number of repetitions,
stimulation speed and decoding performance must be carefully balanced.
Performant classifiers can shift this balance to the benefit of \ac{itr}.
Several techniques are suited to decode~\acp{erp}.
Below are some methods that can achieve state-of-the art performance and are
relevant to this work.

### 2024-10-07 chapters/conclusion.tex
cover page
funding

\todo{https://sci-hub.ru/https://ieeexplore.ieee.org/abstract/document/1642754}
\todo{https://www.sciencedirect.com/science/article/abs/pii/B9780444639349000020}
\todo{BCI principles and practice chapter 19}
\todo{https://www.nature.com/articles/s44222-024-00239-5}

@@

\subsection{Working \emph{with} individuals \acs{sspi}}

@@

\subsection{Working with patients}
gaze independent visual, can be good current existing solution
take into accounrecommendations of \textcite{FriedOken2020}

>>>>

\chapter{Conclusion \& recommendations}
\todo{https://sci-hub.ru/https://ieeexplore.ieee.org/abstract/document/1642754}
\todo{https://www.sciencedirect.com/science/article/abs/pii/B9780444639349000020}
\todo{BCI principles and practice chapter 19}
\todo{https://www.nature.com/articles/s44222-024-00239-5}

@@

flexible models

kron-sum is more flexible than kron, tlda even more flexible

similarly, bttda is more flexible than hoda

regularization is absolutely necessary, and when faced with few data relative
to dimensionality, good first instinct to pick a simple, restrictive structure.

neural data is not linear, everything is fuzzy
when optimizing for performance, it will soon become clear that there are
settings where these assumptions can be broken to better describe the ongoing
ineractions.
if possible, in doing this, embue as much with prior information about the problem

\subsection{Working \emph{with} individuals with \acs{sspi}}

@@

gaze independent visual, can be good current existing solution
take into accounrecommendations of \textcite{FriedOken2020}

### 2024-10-07 chapters/covert_align.tex
Started patients paper
Comments hakim
Comments Marc
Update bttda formulation
cover page
funding

\emph{This chapter has been published as part of~\textcite{VanDenKerchove2024}}

\section{Introduction}
\todo{mention cble and wcble in methods}

\textcite{Arico2014} observed higher variability in single-trial P3 peak
latencies relative to stimulus onset during covert VSA compared to overt VSA.
This latency variability contributes to reduced covert VSA decoding performance.
While they proposed an analysis and performance prediction method, they did not provide a
decoding solution.
They suggested that compensating for latency jitter could enhance covert VSA
decoding, but did not actually verify this hypothesis directly.
Additionally, \cite{Hardiansyah2020} developed a classifier for
covert VSA ERPs, exploiting single-trial latency features in combination with
amplitude features for classification with a support vector machine.
They demonstrated the positive influence of single-trial ERP component latency
features on covert VSA inference, yet did not attempt to correct the amplitude
features for these latencies.
\todo{Should we already mention Frenzel here?}
\textcite{Frenzel2011} introduced a similar protocol to our split VSA
setting.
They showed that it is possible to perform split VSA and that, in this
case, visuospatial attention and gaze direction can separately be decoded using
classical ERP techniques.
To the best of our
knowledge, this is the sole study that investigated split attention in ERP-based BCIs.
However, \textcite{Frenzel2011} considered their interface only for the case
where the user actively intends to select both targets determined by the gaze
and the VSA.
In the split VSA setting considered in our work, we rather instruct the participant to ignore
the distractor and only attend the cued target, since we are interested in
decoding the visuospatial attention only.
\todomvh{avoid overlap. it seems already explained in previous sentences. The reader can find details in
  [6].}

frenzel
\todo{part about split attention etc from chapter on covert alignment to here}

@@

\subsection{Data collection}
We evaluate our approach on the
publicly available BNCI2014-009 dataset~\cite{Aloise2012a} and one specifically recorded for this study, designed to
probe different modalities of covert VSA.
\todomvh{first explain BNCI2014-009 using fig. 6.1 or revert the announcement
in the previous sentence.}

\subsubsection{CVSA-ERP dataset}

@@

	\caption{
		\subref{fig:interface} Interfaces and visuospatial attention (VSA)
    conditions in the CVSA-ERP and BNCI2014-009 datasets.
		In the CVSA-ERP oddball BCI interface, screen targets are intensified one after
    the other in pseudorandom order while the participant can either pay overt,
    covert, or split VSA to the cued target.
		In the BNCI2014-009 overt VSA interface, entire rows and columns are
		intensified at once. In its covert counterpart,
		groups of 6 letters are intensified one after the other, partly relying on
    feature attention.
		\subref{fig:erps} Contrast between target (color) and non-target, and
    distractor (gray) and non-target grand average event-related potentials per VSA condition and dataset.
    Overt VSA yields a strong modulation of the N1 component in both datasets;
    the P3 amplitude decreases with the degree of split VSA.
    In split VSA, N1 and P2 are more prominently evoked by the distractor,
    while the P3 is evoked by the target.
	}
\end{figure*}

@@

Participants signed the informed consent form and were seated at a distance of
60 cm before a CRT-emulating monitor (VPixx
Technologies, Canada) operating at a refresh rate of 120Hz, displaying 6
circular white targets with a diameter of 4.15° visual angle and laid out on a hexagon
with a radius of 12.28° of visual angle centered on the midpoint of the screen,
conforming to the interface proposed by \textcite{Treder2010}
(Figure~\ref{fig:layout_targets})\todo{figure interface in four panels and
update all refs below}.
A hexagonal layout interface with an empty center and a low number of targets
counteracts target crowding and, as long as the subject’s gaze is within the hexagon of
targets, no other target can be between the subject’s gaze and a covertly
attended target.
Targets are full-contrast white and were intensified by scaling them to a
larger size (5.60° of visual angle, Figure~\ref{fig:layout_intense}) instead of changing the contrast to avoid Troxler-fading\footnote{The optical illusion of disappearing unchanging stimuli
experienced when visually fixating~\cite{Troxler1804}.}~\cite{Treder2010} in the
peripheral visual field.
Stimuli were presented using PsychopPy (version 2023.1.3)~\cite{Peirce2019}.
\todo{figures}

The participant was instructed to press the space bar when ready for a block
of stimulations.
Then, one target was indicated as the cue and the participant
was instructed to count the number of intensifications of the cued target
during the following block of stimulations.
After pressing the space bar again, a blue crosshair appeared, and the subject
was instructed to fixate their gaze on the blue crosshair for the duration of
the stimulation block (Figure~\ref{fig:layout_cross_overt} and Figure~\ref{fig:layout_cross_covert}).
The position of this crosshair determined the VSA condition for this trial:
overt VSA when the crosshair was at the same location as the cued target,
covert VSA when the crosshair appeared in the center of the screen, and split
VSA when the crosshair appeared on a different target than the cued one.
After pressing the space bar again and a delay of 5 seconds, the stimulation block
starts.

@@

\subsubsection{Decoders}
We will compare \ac{wcble} and \ac{cble} as described in
Section~\ref{sec:wcble/methods}, with their base classifier and tLDA and with
Riemannian Geometry
approaches that rely on spatial covariance as features.
Together with tLDA, Riemannian Geometry generally reaches state-of-the-art decoding
performance~\cite{Lotte2018}.
We implemented two Riemannian Geometry pipelines.
The first one estimates shrunk covariances from the ERPs filtered with 6 XDAWN
filters, projects these covariances to a tangent space and classifies the
result using $L_2$-regularized logistic regression
(XDAWNCov-TS-LR)~\cite{Cecotti2017}.
Secondly, we adopt the pipeline from \cite{Aydarkhanov2020}, since their work shows
favorable performance in the presence of single-trial ERP latency jitter.
Shrunk spatial covariance matrices are estimated from epochs that are
augmented by concatenating the average target and average non-target ERP as
extra channels, projected to tangent space and classified using
$L_2$-regularized logistic regression (ERPCov-TS-LR).

@@

	\begin{subfigure}{\linewidth}
		\caption{}
		\label{fig:block_evaluation}
		\includegraphics[width=\linewidth]{figures/covert_align/figure4b.pdf}
	\end{subfigure}
	\caption{
		\subref{fig:single_trial_roc_auc_diff} Difference in cross-validated
    single-trial classification receiver operating characteristic curve ($\Delta$ROC-AUC)
    between Classifier-based Latency Estimation (CBLE), Woody
    CBLE (WCBLE),	and their first-stage classifier (tLDA).
    95\% confidence intervals were determined using $k=1000$ bootstrapping.
    Our proposed WCBLE decoder outperforms tLDA and CBLE for covert and split
		($d=3$)	visuospatial attention (VSA). CBLE scores on par with tLDA.
		\subref{fig:block_evaluation} Cross-validated target selection accuracy for
    all decoders plotted n function of the number of test repetitions in
    different VSA conditions. Significance was determined using one-sided
    (WCBLE $>$ other) Wilcoxon rank-sum tests using False Discovery Rate correction
    ($*= p<0.05$, $**=p<0.01$, $***=p<0.001$). WCBLE generally achieves highest
    covert VSA target selection accuracy.}
\end{figure}
Wilcoxon signed-rank tests controlled for multiple comparisons by Benjamini
and Hochberg's False Discovery Rate procedure (FDR) revealed that
for the BNCI2014-009 dataset, WCBLE significantly outperformed tLDA
($\Delta\mathrm{ROC-AUC} = 0.019$, $p=0.004$) and CBLE
($\Delta\mathrm{ROC-AUC} = 0.016$, $p=0.036$) for covert VSA
but was significantly outperformed by tLDA in overt VSA decoding
($\Delta\mathrm{ROC-AUC}=-0.004$, $p=0.040$).
For the CVSA-ERP dataset, WCBLE
also achieved significantly better covert VSA performance than tLDA ($\Delta\mathrm{ROC-AUC}
	= 0.023$, $p=0.041$) and CBLE ($\Delta\mathrm{ROC-AUC}
	= 0.036$, $p=0.024$) .
We found no significant difference in WCBLE performance over tLDA in the split
VSA conditions in the CVSA-ERP dataset, but results show a clear trend of
increase in WCBLE performance over tLDA and CBLE as $d$ increases
CBLE failed to significantly outperform its first-stage classifier tLDA in all
evaluated VSA conditions.
Table~\ref{tab:single_trial_full} reports all single-trial classification scores for all considered models, datasets
and conditions.
\todo{table?}

@@

	\caption{
		\subref{fig:covert_cross_eval},\subref{fig:aloise2012_cross_eval}
    Difference in cross-validated area under the receiver
		operating characteristic curves between Classifier-based Latency Estimation
    (CBLE) and Woody CBLE (WCBLE) across conditions for the CVSA-ERP and
    BNCI2014-009 datasets respectively.
    A decoder is each	time trained on a visuospatial attention (VSA) condition
    and tested on all VSA conditions.
    WCBLE yields an improvement in most
		non-overt VSA settings, indicating it is more invariant to eye gaze than
    CBLE and tLDA.
		\subref{fig:jitter} Jitter characterized as the inter-quartile range (IQR)
		of target epochs for different VSA conditions. Overt VSA exhibits lower
    jitter than other conditions.
    Significance of differences
    was determined with two-sided Wilcoxon rank-sum tests with False Discovery
    Rate correction on per-subject jitter ($*= p<0.05$, $**=p<0.01$,
    $***=p<0.001$, $****=p<0.0001$).
	}
\end{figure*}
When comparing CBLE and tLDA, we do not observe large differences in any of the
evaluated settings, similar to the within-subject conditions.
On the contrary, when considering the comparisons between WCBLE and tLDA, we
see that performance is on par or greater using WCBLE for most conditions,
except for within-overt VSA decoding.

>>>>

\chapter{Compensating \acs{erp} jitter for gaze-independence}
\label{sec:covert-align}
\emph{This chapter has been published as part of~\textcite{VanDenKerchove2024}}

\section{Introduction}

The decoding methods described in \autoref{sec:stbf-struct} and
\autoref{sec:bttda} are rather general attempts at increasing \ac{erp} decoding accuracy.
Given the problem statement of this work, we are also interested in targeted
methods that would specifically advance the development of gaze-independent
visuall \acp{bci} (see \label{sec:gaze-independence/sota}).
Methods of directly enhancing gaze-independent decoding through optimizing
decoder design have not been widely studied (see
\label{sec:gaze-independence/sota/decoding}).
We assume that advances in gaze-independent decoding can be made if properties
of covert \ac{vsa} that affect performance can be identified and, in some way,
accounted for.
The latency estimation and alignment approach proposed in \autoref{sec:wcble}
was specifically designed for this purpose.

\textcite{Arico2014} observed higher variability in single-trial P3 peak
latencies relative to stimulus onset during covert \ac{vsa} compared to overt
\ac{vsa}.
This latency variability contributes to reduced covert VSA decoding performance.
While they proposed an analysis and performance prediction method, they did not provide a
decoding solution.
They suggested that compensating for latency jitter could enhance covert VSA
decoding, but did not actually verify this hypothesis directly.
Additionally, \cite{Hardiansyah2020} developed a classifier for
covert VSA ERPs, exploiting single-trial latency features in combination with
amplitude features for classification with a support vector machine.
They demonstrated the positive influence of single-trial ERP component latency
features on covert VSA inference, yet did not attempt to correct the amplitude
features for these latencies.

Furthermore, as mentioned in~\autoref{sec:gaze-independence/sota/visual} and
\autoref{sec:gaze-independence/approach/cvsa-erp},
our goal is also to explore more flexible gaze-independent settings that the
usual covert \ac{vsa} with central gaze fixation reported in most literature.
\textcite{Frenzel2011} introduced a similar protocol to the proposed split
\ac{vsa} setting (see~\autoref{fig:gaze/vsa/free}).
They showed that it is possible to perform split VSA and that, in this
case, visuospatial attention and gaze direction can separately be decoded using
classical ERP techniques.
To the best of our
knowledge, this is the sole study that investigated split attention in ERP-based BCIs.
However, \textcite{Frenzel2011} considered their interface only for the case
where the user actively intends to select both targets determined by the gaze
and the VSA.
In the split VSA setting considered in our work, we rather instruct the participant to ignore
the distractor and only attend the cued target, since we are interested in
decoding the visuospatial attention only.

@@

\subsection{Data collection}
We evaluate our approach on a dataset specifically recorded for this study, designed to
probe different modalities of covert VSA, and on the publicly available
BNCI2014-009 dataset~\cite{Aloise2012a}.

\subsubsection{CVSA-ERP dataset}
\label{sec:covert-align/stimulation}

@@

  \caption[Stimulation interfaces and evoked \acp{erp}.]{
		\subref{fig:interface} Interfaces and visuospatial attention (VSA)
    conditions in the CVSA-ERP and BNCI2014-009 datasets.
		In the CVSA-ERP oddball BCI interface, screen targets are intensified one after
    the other in pseudorandom order while the participant can either pay overt,
    covert, or split VSA to the cued target.
		In the BNCI2014-009 overt VSA interface, entire rows and columns are
		intensified at once. In its covert counterpart,
		groups of 6 letters are intensified one after the other, partly relying on
    feature attention.
		\subref{fig:erps} Contrast between target (color) and non-target, and
    distractor (gray) and non-target grand average event-related potentials per VSA condition and dataset.
    Overt VSA yields a strong modulation of the N1 component in both datasets;
    the P3 amplitude decreases with the degree of split VSA.
    In split VSA, N1 and P2 are more prominently evoked by the distractor,
    while the P3 is evoked by the target.
	}
\end{figure*}

@@

Participants signed the informed consent form and were seated at a distance of
60 cm before a CRT-emulating monitor (VPixx
Technologies, Canada) operating at a refresh rate of 120Hz, displaying 6
circular white targets with a diameter of 4.15° visual angle and laid out on a hexagon
with a radius of 12.28° of visual angle centered on the midpoint of the screen,
conforming to the interface proposed by \textcite{Treder2010}
(\autoref{fig:covert-align/interface/interface})
A hexagonal layout interface with an empty center and a low number of targets
counteracts target crowding and, as long as the subject’s gaze is within the hexagon of
targets, no other target can be between the subject’s gaze and a covertly
attended target.
Targets are full-contrast white and were intensified by scaling them to a
larger size (5.60° of visual angle,
(\autoref{fig:covert-align/interface/intensification}) instead of changing the contrast to avoid Troxler-fading\footnote{The optical illusion of disappearing unchanging stimuli
experienced when visually fixating~\cite{Troxler1804}.}~\cite{Treder2010} in the
peripheral visual field.
Stimuli were presented using PsychopPy (version 2023.1.3)~\cite{Peirce2019}.

\begin{figure}
  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/covert_align/interface_1.tikz.tex}
    \caption[The stimulation interface.]{The stimulation interface based on the visual Hex-o-Spell
    \acs{bci}~\cite{Treder2010}.}
    \label{fig:covert-align/interface/interface}
  \end{subfigure}\hfill%
  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/covert_align/interface_2.tikz.tex}
    \caption[Intensification.]{When a target is intensified, it is enlarged for
    a very short time.}
    \label{fig:covert-align/interface/intensification}
  \end{subfigure}

  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/covert_align/interface_3.tikz.tex}
    \caption[Gaze fixated on a target.]{In the overt and split \ac{vsa}
    settings, the
    participant is cued to fixate on one of the targets.}
    \label{fig:covert-align/interface/overt}
  \end{subfigure}\hfill%
  \begin{subfigure}[t]{.45\textwidth}
    \input{figures/covert_align/interface_4.tikz.tex}
    \caption[Gaze fixated centrally]{In the covert \ac{vsa} setting, the participant is
    cued to fixate the center of the interface.}
    \label{fig:covert-align/interface/covert}
  \end{subfigure}
\end{figure}

The participant was instructed to press the space bar when ready for a block
of stimulations.
Then, one target was indicated as the cue and the participant
was instructed to count the number of intensifications of the cued target
during the following block of stimulations.
After pressing the space bar again, a blue crosshair appeared, and the subject
was instructed to fixate their gaze on the blue crosshair for the duration of
the stimulation block (\autoref{fig:covert-align/interface/overt} and
\autoref{fig:covert-align/interface/overt}).
The position of this crosshair determined the VSA condition for this trial:
overt VSA when the crosshair was at the same location as the cued target,
covert VSA when the crosshair appeared in the center of the screen, and split
VSA when the crosshair appeared on a different target than the cued one.
After pressing the space bar again and a delay of 5 seconds, the stimulation block
starts.

@@

\subsubsection{Decoders}
We will compare \ac{wcble} and \ac{cble} as described in
\autoref{sec:wcble/methods}, with their base classifier and tLDA and with
Riemannian Geometry
approaches that rely on spatial covariance as features.
Together with tLDA, Riemannian Geometry generally reaches state-of-the-art decoding
performance~\cite{Lotte2018}.
We implemented two Riemannian Geometry pipelines.
The first one estimates shrunk covariances from the ERPs filtered with 6 XDAWN
filters, projects these covariances to a tangent space and classifies the
result using $L_2$-regularized logistic regression
(XDAWNCov-TS-LR)~\cite{Cecotti2017}.
Secondly, we adopt the pipeline from \cite{Aydarkhanov2020}, since their work shows
favorable performance in the presence of single-trial ERP latency jitter.
Shrunk spatial covariance matrices are estimated from epochs that are
augmented by concatenating the average target and average non-target ERP as
extra channels, projected to tangent space and classified using
$L_2$-regularized logistic regression (ERPCov-TS-LR).

@@

	\begin{subfigure}{\linewidth}
		\caption{}
		\label{fig:block_evaluation}
		\includegraphics[width=\linewidth]{figures/covert_align/figure4b.pdf}
	\end{subfigure}
  \caption[Decoder performance.]{%
		\subref{fig:single_trial_roc_auc_diff} Difference in cross-validated
    single-trial classification receiver operating characteristic curve ($\Delta$ROC-AUC)
    between Classifier-based Latency Estimation (CBLE), Woody
    CBLE (WCBLE),	and their first-stage classifier (tLDA).
    95\% confidence intervals were determined using $k=1000$ bootstrapping.
    Our proposed WCBLE decoder outperforms tLDA and CBLE for covert and split
		($d=3$)	visuospatial attention (VSA). CBLE scores on par with tLDA.
		\subref{fig:block_evaluation} Cross-validated target selection accuracy for
    all decoders plotted n function of the number of test repetitions in
    different VSA conditions. Significance was determined using one-sided
    (WCBLE $>$ other) Wilcoxon rank-sum tests using False Discovery Rate correction
    ($*= p<0.05$, $**=p<0.01$, $***=p<0.001$). WCBLE generally achieves highest
    covert VSA target selection accuracy.}
\end{figure}
Wilcoxon signed-rank tests controlled for multiple comparisons by Benjamini
and Hochberg's False Discovery Rate procedure (FDR) revealed that
for the BNCI2014-009 dataset, WCBLE significantly outperformed tLDA
($\Delta\mathrm{ROC-AUC} = 0.019$, $p=0.004$) and CBLE
($\Delta\mathrm{ROC-AUC} = 0.016$, $p=0.036$) for covert VSA
but was significantly outperformed by tLDA in overt VSA decoding
($\Delta\mathrm{ROC-AUC}=-0.004$, $p=0.040$).
For the CVSA-ERP dataset, WCBLE
also achieved significantly better covert VSA performance than tLDA ($\Delta\mathrm{ROC-AUC}
	= 0.023$, $p=0.041$) and CBLE ($\Delta\mathrm{ROC-AUC}
	= 0.036$, $p=0.024$) .
We found no significant difference in WCBLE performance over tLDA in the split
VSA conditions in the CVSA-ERP dataset, but results show a clear trend of
increase in WCBLE performance over tLDA and CBLE as $d$ increases
CBLE failed to significantly outperform its first-stage classifier tLDA in all
evaluated VSA conditions.
Table~\ref{tab:covert-align/results/single-trial} reports all single-trial classification scores for all considered models, datasets
and conditions.

\begin{table}
  \centering
  \makebox[\textwidth][c]{%
    \input{chapters/covert_align_performance.tex}
  }
  \caption{Cross-validated single-trial classification area under the receiver operating characteristic
curve for all evaluated models, visuospatial attention conditions and
  datasets.}
  \label{tab:covert-align/results/single-trial}
\end{table}

@@

  \caption[Cross-condition classifier performance and estimated jitter.]{
		\subref{fig:covert_cross_eval},\subref{fig:aloise2012_cross_eval}
    Difference in cross-validated area under the receiver
		operating characteristic curves between Classifier-based Latency Estimation
    (CBLE) and Woody CBLE (WCBLE) across conditions for the CVSA-ERP and
    BNCI2014-009 datasets respectively.
    A decoder is each	time trained on a visuospatial attention (VSA) condition
    and tested on all VSA conditions.
    WCBLE yields an improvement in most
		non-overt VSA settings, indicating it is more invariant to eye gaze than
    CBLE and tLDA.
		\subref{fig:jitter} Jitter characterized as the inter-quartile range (IQR)
		of target epochs for different VSA conditions. Overt VSA exhibits lower
    jitter than other conditions.
    Significance of differences
    was determined with two-sided Wilcoxon rank-sum tests with False Discovery
    Rate correction on per-subject jitter ($*= p<0.05$, $**=p<0.01$,
    $***=p<0.001$, $****=p<0.0001$).
	}
\end{figure*}
When comparing CBLE and tLDA, we do not observe large differences in any of the
evaluated settings, similar to the within-subject conditions.
On the contrary, when considering the comparisons between WCBLE and tLDA, we
see that performance is on par or greater using WCBLE for most conditions,
except for within-overt VSA decoding.

### 2024-10-07 chapters/funding.tex
funding

\todo{acknowledgement wcble paper}
We thank the Flemish Supercomputer Center (VSC) and the High-Performance
Computing (HPC) center of the KU Leuven for allowing us to execute our
computational experiments on their systems.
We also wish to acknowledge dr.\ Axel Faes for his inspiration in conceptualizing
the work presented in Chapter~\ref{sec:bttda}.

>>>>

Arne Van Den Kerchove is supported by the special research fund of the KU Leuven
(GPUDL/20/031).
prof. Marc Van Hulle is supported by research grants received from
the European Union’s Horizon Europe Marie Sklodowska-Curie Action program
(grant agreement No. 101118964), the European Union’s Horizon 2020 research and
innovation program (grant agreement No. 857375), the special research fund of
the KU Leuven (C24/18/098), the Belgian Fund for Scientific Research – Flanders
(G0A4118N, G0A4321N, G0C1522N), and the Hercules
Foundation (AKUL 043).

We thank the Flemish Supercomputer Center (VSC) and the High-Performance
Computing (HPC) center of the KU Leuven for allowing us to execute our
computational experiments on their systems.

We also thank,  Arno Libert,
Tjaša Mlinarič, Yide Li and Mani Mirsaeedi for their help in data collection,
and in general the colleagues at the KU Leuven Computational Neurosciences
group at KU Leuven and the BCI team at the University of Lille.
Furthermore, we thank the support teams at TRAINM Neuro Rehab Clinics Antwerp,
the UZ Leuven Neuromuscular Reference Renter and the Fondation Partage et Vie.
Finally, we wish to acknowledge dr.\ Axel Faes for his inspiration in conceptualizing
the work presented in Chapter~\ref{sec:bttda}.
